---
title: "Reproducible Research: Peer Assessment 1 by Alex Konkel"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we'll need to unzip the file to make the data accessible.
``` {r}
walk <- unzip('activity.zip')
```

And now we'll need to get the file loaded for analysis.
I'll also convert the date column to the date class.

``` {r}
walk <- read.csv('activity.csv',header=TRUE,as.is=TRUE)
walk$date=as.Date(walk$date,format='%Y-%m-%d')
```
And now we should be good to go.

## What is mean total number of steps taken per day?
The instructions say that we can ignore the NAs, so the total number of steps per day  
is pretty straightforward.
``` {r}
totPerDay <- tapply(walk$steps,walk$date,sum,na.rm=TRUE)
totPerDay
```

And how about a histogram of that?

```{r,fig.height=4}
hist(totPerDay,breaks=20,col='light blue')
```

We can see that no recorded steps is about as frequent as the other most common bin (~10000 steps).
Finally, we want the mean and medium across days.  That's pretty similar to the total.
``` {r}
meanSteps <- mean(totPerDay)
medianSteps <- median(totPerDay)
```

The mean is `r round(meanSteps,2)` and the median is `r medianSteps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
