---
title: "Reproducible Research: Peer Assessment 1 by Alex Konkel"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we'll need to unzip the file to make the data accessible.
``` {r}
walk <- unzip('activity.zip')
```

And now we'll need to get the file loaded for analysis.
I'll also convert the date column to the date class.

``` {r}
options(scipen=1, digits=2)
walk <- read.csv('activity.csv',header=TRUE,as.is=TRUE)
walk$date <- as.Date(walk$date,format='%Y-%m-%d')
```
And now we should be good to go.  The options call will affect how some numbers display later.

## What is mean total number of steps taken per day?
The instructions say that we can ignore the NAs, so the total number of steps per day  
is pretty straightforward.
``` {r}
totPerDay <- tapply(walk$steps,walk$date,sum,na.rm=TRUE)
totPerDay
```

And how about a histogram of that?

```{r,fig.height=4}
hist(totPerDay,breaks=20,col='light blue',xlab='Total Steps per Day',main='Histogram of Steps per Day')
```

We can see that no recorded steps is about as frequent as the other most common bin (~10000 steps).
Finally, we want the mean and medium across days.  That's easy.
``` {r}
meanSteps <- mean(totPerDay)
medianSteps <- median(totPerDay)
```

The mean is `r meanSteps` and the median is `r medianSteps`.

## What is the average daily activity pattern?
Next, instead of looking at the level of days we'll look at the level of 5 minute intervals within days.  
Here's a plot of the average number of steps per interval.
``` {r, fig.height = 4}
totPerInc <- tapply(walk$steps,walk$interval,mean, na.rm=TRUE)
plot(totPerInc,type='l',xlab='Interval of the Day',ylab='Average Steps per Interval')
maxInc <- which.max(totPerInc)
maxSteps <- totPerInc[maxInc]

```
The most steps, on average, occur during the `r maxInc`th interval when an average of `r maxSteps` steps are taken.  
Since the intervals are 5 minutes long, this occurs `r 5*maxInc` minutes into the day, or `r strftime( as.POSIXct( "1970-01-01" ) + as.difftime( 5*maxInc, units="mins"), "%H:%M")` "military time".

## Imputing missing values
Step 1 for filling in the missing values is to determine how many there are.
``` {r}
totMissing <- length(which(is.na(walk$steps)))
```
There are `r totMissing` missing data points.

To fill in the missing points, I'm going to look ahead to the next part of the assignment a bit.
Since some days are completely missing, I don't want to just use the average within a day, because that would still be unknown.
Instead, I'm going to use a regression to predict the number of steps by the interval, day, and day of the week.
``` {r}
walk$day <- weekdays(walk$date)
mod1 <- lm(steps ~ date+interval+day,data=walk)
missPoints <- subset(walk,is.na(steps))
predictions <- predict(mod1,new=missPoints[,2:4])
missPoints$steps <- predictions
```
So now we have the missing observations separated out and replaced with an imputed guess.  The model isn't great;  
the adjusted R squared is less than .01.  But the regression says that some days are predictive  
as is the interval in the day, so hopefully the imputation is better than a very simple average.

Now we want to make a new dataset that incorporates the imputation and look at the mean and median steps across days.  
To aid comparison to the original dataset, I'll put the earlier histogram here on the bottom.
``` {r, fig.height=4}
require(dplyr)
walk2 <- walk %>% filter(!is.na(steps)) %>% bind_rows(missPoints) %>% arrange(date,interval)
totPerDay2 <- tapply(walk2$steps,walk2$date,sum,na.rm=TRUE)
par(mfrow=c(2,1),mar=c(4,4,2,0))
hist(totPerDay2,breaks=20,col='light blue',xlab='Total Steps per Day, With Imputation',main='',cex.axis=.6,cex.lab=.6,ylim=c(0,12))
hist(totPerDay,breaks=20,col='light blue',xlab='Total Steps per Day, Original',main='',cex.axis=.6,cex.lab=.6,ylim=c(0,12))
meanSteps2 <- round(mean(totPerDay2),2)
medianSteps2 <- round(median(totPerDay2),2)
```
After imputaton, the mean is `r meanSteps2` and the median is `r medianSteps2`.  
This is a difference of `r meanSteps2-meanSteps` for the average and `r medianSteps2-medianSteps` for the median,
which is a noticeable difference.   We can also see a shift in the histogram, presumably because days with 0 steps due to NAs now have some.

## Are there differences in activity patterns between weekdays and weekends?
Finally, we want to compare weekend days and weekdays.  I already labeled the days of the week earlier, 
so now we just have to compress those into a new factor variable.  
``` {r}
walk2 <- walk2 %>% mutate(day2 = ifelse(day=='Saturday' | day=='Sunday','weekend','weekday')) %>%
  mutate(day2 = as.factor(day2))
```
Now we want a line plot of the average number of steps per interval, separated for weekends and weekdays.
``` {r, fig.height=4}
require(ggplot2)
walk3 <- walk2 %>% group_by(interval,day2) %>% summarize(average=mean(steps))
ggplot(walk3,aes(x=interval,y=average))+geom_line()+facet_grid(day2~.)+labs(x='Interval of the Day', y='Average Steps Taken')
```

It looks like weekdays have a more prolonged build-up before the morning spike, and that spike is more consistent (there are two spikes on weekends).  More steps happen during the middle of the day on weekends, but the end of the day/evening appear fairly similar.